---
title: "Exxon Marketwatch Analysis"
author: "Trist'n Joseph"
date: "7/25/2020"
output: html_document
---

Packages
```{r}
# Package to assist with article scarping 
library(rvest) 

# Packages to assist with text cleaning 
library(stringr)
library(tm)
library(tidytext)

# Packages to assit with polarity calculation
library(rJava)
library(qdap)

# Package to assit with data manipulation & plotting 
library(tidyverse)
library(ggthemes)

# Package to collect financial data 
library(quantmod)

```

IGNORE 
```{r}
exxon_articles$Date <- as.character(exxon_articles$Date)
exxon_articles$Date <- parse_date(exxon_articles$Date, format ="%m/%d/%Y")
exxon_articles$Article <- as.character(exxon_articles$Article)

head(exxon_articles)

exxon_articles%>%
  select(Date)%>%
  arrange(Date)


anyNA(exxon_articles$Article)
sum(is.na(exxon_articles))
exxon_articles$Article[c(6,144)]

exxon_articles <- exxon_articles[-c(6,144), ]
sum(is.na(exxon_articles))

exxon_articles$Article[31]

exxon_articles_cleaned <- ExxonArticle_2
exxon_articles_cleaned$Article <- tolower(exxon_articles_cleaned$Article)
exxon_articles_cleaned$Article <- removePunctuation(exxon_articles_cleaned$Article)
exxon_articles_cleaned$Article <- removeSymbols(exxon_articles_cleaned$Article)
exxon_articles_cleaned$Article <- removeWords(exxon_articles_cleaned$Article, words = c(stopwords(kind = "en"), "marketwatch", "mw", "zachs", "zacks"))
exxon_articles_cleaned$Article <- stripWhitespace(exxon_articles_cleaned$Article)


```

Exxon Article Data -- Local File
```{r}

exxon_articles <- read_csv("E:/Projects/Sentiment Analysis/ExxonArticle_2.csv")

# Removes unnecessary columns from the working df
exxon_articles$X1 <- NULL
exxon_articles$Date.X <- NULL
ExxonArticle_2 <- exxon_articles # Renamed to match Tyler's file
rm(exxon_articles)

# Renames columns in the working df
names(ExxonArticle_2) <- c("Date", "Article")

ExxonArticle_2 <- ExxonArticle_2 %>%
  arrange(Date)

ExxonArticle_Analysis <- ExxonArticle_2 %>%
  filter(Date >= "2020-03-01" & Date <= "2020-06-20")
rm(ExxonArticle_2)

# Calculates the polarity of each article
Polarity <- polarity(ExxonArticle_Analysis$Article) 

# Adds the polarity to the working df
ExxonArticle_Analysis$Polarity <- Polarity$all$polarity
rm(Polarity)

# Aggregates the polarity by day since each day can have multiple articles & multiple polarities
ExxonArticle_TotalPolarity <- ExxonArticle_Analysis %>%
  group_by(Date) %>%
  summarise(Total_polarity = sum(Polarity, na.rm = T)) %>%
  arrange(Date)

ExxonArticle_TotalPolarity$Date
anyNA(ExxonArticle_TotalPolarity$Total_polarity)

```

Exxon Financial Data
```{r}

startDate = as.Date("2020-03-4")
endDate = as.Date("2020-06-21")
# DF is automatically saved as the ticker symbol
getSymbols("XOM", from =startDate, to = endDate)

# Extracts the price column as a numeric list from an xts object 
XOM_Price <- as.numeric(XOM$XOM.Close)
# Drops dates where no articles where recorded
XOM_Price <- XOM_Price[-c(2,3,18,21,75)] 

# Extracts the volume column as a numeric list from an xts object 
XOM_Volume <- as.numeric(XOM$XOM.Volume)
# Drops dates where no articles where recorded
XOM_Volume <- XOM_Volume[-c(2,3,18,21,75)] 

rm(XOM)

```

Data Preparation for Models
```{r}
# Combine the working df with price and volume lists 
Exxon_Regression <- cbind(ExxonArticle_TotalPolarity, XOM_Price, XOM_Volume)
rm(ExxonArticle_TotalPolarity, XOM_Price, XOM_Volume, ExxonArticle_Analysis)

# Raname columns 
colnames(Exxon_Regression) <- c("Date", "Polarity", "Price", "Volume")

# Create new variable to record observations as 'Positive' if polarity >= 0
Exxon_Regression$Polarity_Sign <- ifelse(Exxon_Regression$Polarity >= 0, "Positive", "Negative")

# Create new variable to find the standard prices of exxon stock
Exxon_Regression <- Exxon_Regression %>%
  mutate(Price_Standard = (Price - mean(Price))/sd(Price))

# Create grouping variable 'Month' which is the recoreded month of each observation
Exxon_Regression$Month <- factor(months(Exxon_Regression$Date), levels = month.name)

#write.csv(Exxon_Regression, "E:/Projects/Sentiment Analysis/ExxonRegression.csv") # sending to Tyler

cols <- c("Polarity" = "red", "Standardized Price" = "blue")
ggplot(data = Exxon_Regression)+
  geom_line(mapping = aes(x = Date, y = Polarity, col = "Polarity"))+
  geom_line(mapping = aes(x = Date, y = Price_Standard, col = "Standardized Price"))+
  scale_x_date(date_breaks = "months", date_labels = "%b-%Y")+
  scale_color_manual(name = "Lines", values = cols)+
  theme_tufte()+
  ylab("Standardized Values")+
  xlab("Date")+
  ggtitle("Relationship Between Polarity & Standardized Price")

```

Regression -- Forecasting -- Function with different weightings
```{r}

predict_prices1 <- function(initial_price, initial_sentiment, sentiment_vector){
  # This function takes an initial price and a list of sentiments (polarities)
  # where the first sentiment is the initial sentiment
  # Positive news as a higher weighting than negative news 
  # The function returns a list of predicted stock prices 
  # Predictions are made based on the recursive relationship defined below 
  
  # Function is optimized using intermediate value theorem and the evaluation metric
  
  stock_price <- c()
  if (initial_price < 0){
    new_stock_price <- initial_price + (0.1*initial_price*initial_sentiment)
  }
  else {
    new_stock_price <- initial_price + (0.105*initial_price*initial_sentiment)
  }
  stock_price <- append(stock_price, new_stock_price, after = length(stock_price))
  for(i in 1:length(sentiment_vector)){
    if(sentiment_vector[i] < 0){
      new_stock_price <- stock_price[i] + (0.1*stock_price[i]*sentiment_vector[i])
      print("Predicting new stock prices...")
    }
    else{
      new_stock_price <- stock_price[i] + (0.105*stock_price[i]*sentiment_vector[i])
      print("Predicting new stock prices...")
    }
    stock_price <- append(stock_price, new_stock_price, after = length(stock_price))
  }
  final_stock_prices <- c(initial_price, stock_price)
  print("Done")
  return(final_stock_prices)
}   
# Constant on sentiment_vector is ideal at 0.1 for negative sentiments & 0.105 for positive sentiments

predicted_prices_1 <- predict_prices1(52.42, -0.07987329, Exxon_Regression$Polarity[2:70])
Exxon_Regression$Predicted_Price_1 <- predicted_prices_1

model1_difference <- Exxon_Regression %>%
  # The average deviation should be minimized for ideal model performance
  select(c(Price, Predicted_Price_1)) %>%
  mutate(Difference = Predicted_Price_1 - Price, Absolute_Difference = abs(Price - Predicted_Price_1)) %>%
  summarize(Avg_Predicted_minus_Actual = mean(Difference), Avg_Absolute_Difference = mean(Absolute_Difference))
model1_difference

model1_difference_by_month <- Exxon_Regression %>%
  # This shows model performance broken down by month
  select(c(Price, Predicted_Price_1, Month)) %>%
  mutate(Difference = Predicted_Price_1 - Price, Absolute_Difference = abs(Price - Predicted_Price_1)) %>%
  group_by(Month) %>%
  summarise(Avg_Predicted_minus_Actual = mean(Difference), Avg_Absolute_Difference = mean(Absolute_Difference))
model1_difference_by_month

cols <- c("Predicted Price" = "red", "Actual Price" = "green")
ggplot(data = Exxon_Regression)+
  geom_line(mapping = aes(x = Date, y = Price, col = "Actual Price"))+
  geom_line(mapping = aes(x = Date, y = Predicted_Price_1, col = "Predicted Price"))+
  scale_x_date(date_breaks = "months", date_labels = "%b-%Y")+
  scale_color_manual(name = "Lines", values = cols) +
  theme_tufte()+
  ylab("Stock Price")+
  xlab("Date")+
  ggtitle("Exxon Mobil Stock Price")

```

Regression -- Forecasting -- Function with different weightings & volume 
```{r}

predict_prices2 <- function(initial_price, initial_sentiment, initial_volume, sentiment_vector, volume_vector){
  # This function takes an initial price, a list of sentiments (polarities), and a list of volumes
  # where the first sentiment is the initial sentiment & first volume is initial volume 
  # Positive news as a higher weighting than negative news 
  # The function returns a list of predicted stock prices 
  # Predictions are made based on the recursive relationship defined below 
  
  # Function is optimized using intermediate value theorem and the evaluation metric 
  
  stock_price <- c()
  if (initial_price < 0){
    new_stock_price <- initial_price + (0.1*initial_price*initial_sentiment) - (0.000000006*initial_volume*initial_sentiment)
  }
  else {
    new_stock_price <- initial_price + (0.105*initial_price*initial_sentiment) - (0.000000006*initial_volume*initial_sentiment)
  }
  stock_price <- append(stock_price, new_stock_price, after = length(stock_price))
  for(i in 1:length(sentiment_vector)){
    if(sentiment_vector[i] < 0){
      new_stock_price <- stock_price[i] + (0.1*stock_price[i]*sentiment_vector[i]) - (0.000000006*volume_vector[i]*sentiment_vector[i])
      print("Predicting new stock prices...")
    }
    else{
      new_stock_price <- stock_price[i] + (0.105*stock_price[i]*sentiment_vector[i]) - (0.000000006*volume_vector[i]*sentiment_vector[i])
      print("Predicting new stock prices...")
    }
    stock_price <- append(stock_price, new_stock_price, after = length(stock_price))
  }
  final_stock_prices <- c(initial_price, stock_price)
  print("Done")
  return(final_stock_prices)
} 
# Constant on volume_vector <= 0.00000001 & >= 0.000000006
# Constant on sentiment_vector is ideal at 0.1 for negative sentiments & 0.105 for positive sentiments
# when constant on volume_vector is within the stated range

predicted_prices_2 <- predict_prices2(52.42, -0.07987329, 24079600, Exxon_Regression$Polarity[2:70], Exxon_Regression$Volume[2:70])
Exxon_Regression$Predicted_Price_2 <- predicted_prices_2

model2_difference <- Exxon_Regression %>%
  # The average deviation should be minimized for ideal model performance
  select(c(Price, Predicted_Price_2)) %>%
  mutate(Difference = Predicted_Price_2 - Price, Absolute_Difference = abs(Price - Predicted_Price_2)) %>%
  summarize(Avg_Predicted_minus_Actual = mean(Difference), Avg_Absolute_Difference = mean(Absolute_Difference))
model2_difference

model2_difference_by_month <- Exxon_Regression %>%
  # This shows model performance broken down by month
  select(c(Price, Predicted_Price_2, Month)) %>%
  mutate(Difference = Predicted_Price_2 - Price, Absolute_Difference = abs(Price - Predicted_Price_2)) %>%
  group_by(Month) %>%
  summarise(Avg_Predicted_minus_Actual = mean(Difference), Avg_Absolute_Difference = mean(Absolute_Difference))
model2_difference_by_month

cols <- c("Predicted Price" = "red", "Actual Price" = "green")
ggplot(data = Exxon_Regression)+
  geom_line(mapping = aes(x = Date, y = Price, col = "Actual Price"))+
  geom_line(mapping = aes(x = Date, y = Predicted_Price_2, col = "Predicted Price"))+
  scale_x_date(date_breaks = "months", date_labels = "%b-%Y")+
  scale_color_manual(name = "Lines", values = cols)+
  theme_tufte()+
  ylab("Stock Price")+
  xlab("Date")+
  ggtitle("Exxon Mobil Stock Price")

```

Regression -- Forecasting -- Function with different weightings & volume (Confidence Interval)
```{r}

create_confidence_intervals <- function(predicted_prices, confidence_level = 0.95){
  # This function creates the confidence intervals for the predicted prices 
  # It accepts a list of prices and an optimal confidence level (default is 95%)
  # Returns a df with confidence interval upper and lower limits 
  
  standard_dev = sd(predicted_prices)
  remainder <- 1 - confidence_level
  confidence_level_probability <- confidence_level + (remainder/2)
  z = qnorm(confidence_level_probability)
  upper_limits <- c()
  for (price in predicted_prices){
    upper_limit <- price + (z*standard_dev/sqrt(length(predicted_prices)))
    upper_limits <- append(upper_limits, upper_limit, after = length(upper_limits))
    print("Calculating upper confidence limits...")
  }
  lower_limits <- c()
  for (price in predicted_prices){
    lower_limit <- price - (z*standard_dev/sqrt(length(predicted_prices)))
    lower_limits <- append(lower_limits, lower_limit, after = length(lower_limits))
    print("Calculating lower confidence limits...")
  }
  df <- cbind(lower_limits, upper_limits)
  df <- as.data.frame(df)
  print("Done")
  return(df)
}

confidence_intervals <- create_confidence_intervals(Exxon_Regression$Predicted_Price_2) # 95%
Exxon_Regression_Price_2 <- Exxon_Regression %>%
  select(c("Date", "Price", "Month", "Predicted_Price_2")) %>%
  mutate(ConfInt_lower=confidence_intervals$lower_limits, ConfInt_upper=confidence_intervals$upper_limits)

confidence_interval_80 <- create_confidence_intervals(Exxon_Regression$Predicted_Price_2, 0.80)
Exxon_Regression_Price_2$ConfInt_lower_80 <- confidence_interval_80$lower_limits
Exxon_Regression_Price_2$ConfInt_upper_80 <- confidence_interval_80$upper_limits

confidence_interval_99 <- create_confidence_intervals(Exxon_Regression$Predicted_Price_2, 0.99)
Exxon_Regression_Price_2$ConfInt_lower_99 <- confidence_interval_99$lower_limits
Exxon_Regression_Price_2$ConfInt_upper_99 <- confidence_interval_99$upper_limits

cols <- c(
  "Predicted Price" = "red",
  "Actual Price" = "green", 
  "95% Confidence Interval"="blue",
  "80% Confidence Interval"="orange",
  "99% Confidence Interval"="dark grey"
  )

ggplot(data = Exxon_Regression_Price_2)+
  geom_line(mapping = aes(x = Date, y = Price, col = "Actual Price"))+
  geom_line(mapping = aes(x = Date, y = Predicted_Price_2, col = "Predicted Price"))+
  # 95 % CI
  #geom_line(mapping = aes(x = Date, y = ConfInt_lower, col ="95% Confidence Interval"), linetype = "dashed")+
  #geom_line(mapping = aes(x = Date, y = ConfInt_upper), col="blue", linetype = "dashed")+
  # 80% CI
  #geom_line(mapping = aes(x = Date, y = ConfInt_lower_80, col ="80% Confidence Interval"), linetype = "dashed")+
  #geom_line(mapping = aes(x = Date, y = ConfInt_upper_80), col="orange", linetype = "dashed")+
  # 99% CI
  geom_line(mapping = aes(x = Date, y = ConfInt_lower_99, col ="99% Confidence Interval"), linetype = "dashed")+
  geom_line(mapping = aes(x = Date, y = ConfInt_upper_99), col="dark grey", linetype = "dashed")+
  scale_x_date(date_breaks = "months", date_labels = "%b-%Y")+
  scale_color_manual(name = "Lines", values = cols)+
  theme_tufte()+
  ylab("Stock Price")+
  xlab("Date")+
  ggtitle("Exxon Mobil Stock Price")

```

Regression -- Forecasting -- Price Difference
```{r}

Exxon_Regression_PriceDiff <- Exxon_Regression[2:nrow(Exxon_Regression), ] %>% 
  select(c(Date, Polarity, Price, Volume, Month))

calculate_darily_price_difference <- function(Prices){
  # This function takes a list of prices
  # Returns a df containing price difference and abs price difference 
  
  row = 1
  Price_Diffs = c()
  while(row < length(Prices)){
    price_diff <- round(Prices[row+1] - Prices[row], 3)
    Price_Diffs <- append(Price_Diffs, price_diff, after = length(Price_Diffs))
    row = row + 1
    print("Calculating price differences...")
  }
  odf <- as.data.frame(Price_Diffs)
  colnames(odf) <- c("True_Price_Difference")
  odf$Abs_Price_Difference <- abs(odf$True_Price_Difference)
  print("Done")
  return(odf)
}  

#Actual_PriceDifferences <- calculate_darily_price_difference(Exxon_Regression$Price)
#Exxon_Regression_PriceDiff <- cbind(Exxon_Regression_PriceDiff, Actual_PriceDifferences) 

cols <- c("Polarity" = "red", "Absolute Price Difference" = "blue")
ggplot(data = Exxon_Regression_PriceDiff)+
  geom_line(mapping = aes(x = Date, y = Abs_Price_Difference, col = "Absolute Price Difference"))+
  geom_line(mapping = aes(x = Date, y = Polarity, col = "Polarity"))+
  scale_x_date(date_breaks = "months", date_labels = "%b-%Y")+
  scale_color_manual(name = "Lines", values = cols)+
  theme_tufte()+
  ylab("")+
  xlab("Date")+
  ggtitle("Relationship Between Price Difference & Polarity of Exxon Mobil Stock")


predict_price_differences <- function(sentiment_vector, volume_vector){
  # This function takes a list of sentiments (polarities) and volumes
  # Returns a df containing predicted price differences and abs predicted price differences 
  # Predictions are solely based on polarity and traded volume 
  
  Price_Diffs <- c()
  n <- 1 
  while(n <= length(sentiment_vector)){
    price_diff <- round(0.00000004*volume_vector[n] - 1.5*sentiment_vector[n])
    Price_Diffs <- append(Price_Diffs, price_diff, after = length(Price_Diffs))
    n = n + 1
    print("Predicting price differences...")
  }
  odf <- as.data.frame(Price_Diffs)
  colnames(odf) <- c("Predicted_Price_Difference")
  odf$Abs_Predicted_Price_Difference <- abs(odf$Predicted_Price_Difference)
  print("Done")
  return(odf)
}
# Constant on volume_vector <= 0.000000045 & >= 0.000000040
# Constant on sentiment_vector is ideal at 1.5 when constant on volume_vector is within the stated range

predicted_price_difference <- predict_price_differences(Exxon_Regression_PriceDiff$Polarity, Exxon_Regression$Volume)
Exxon_Regression_PriceDiff$Predicted_Price_Difference <- predicted_price_difference$Predicted_Price_Difference
Exxon_Regression_PriceDiff$Abs_Predicted_Price_Difference <- predicted_price_difference$Abs_Predicted_Price_Difference

price_difference_model_performance <- Exxon_Regression_PriceDiff %>% 
  # The average deviation should be minimized for ideal model performance
  select(c(Abs_Price_Difference, Abs_Predicted_Price_Difference)) %>% 
  mutate(Deviation = Abs_Predicted_Price_Difference - Abs_Price_Difference) %>% 
  summarise(Average_Deviation = mean(Deviation))
price_difference_model_performance 
 

price_difference_model_performance_by_month <- Exxon_Regression_PriceDiff %>% 
  # This shows model performance broken down by month
  select(c(Abs_Price_Difference, Abs_Predicted_Price_Difference, Month)) %>% 
  group_by(Month) %>% 
  mutate(Deviation = Abs_Predicted_Price_Difference - Abs_Price_Difference) %>% 
  summarise(Average_Deviation = mean(Deviation))
price_difference_model_performance_by_month

cols <- c("Predicted Absolute Price Difference" = "red", "Actual Absolute Price Difference" = "green")
ggplot(data = Exxon_Regression_PriceDiff) +
  geom_line(mapping = aes(x = Date, y = Abs_Price_Difference, col = "Actual Absolute Price Difference"))+
  geom_line(mapping = aes(x = Date, y = Abs_Predicted_Price_Difference, col = "Predicted Absolute Price Difference"))+
  scale_x_date(date_breaks = "months", date_labels = "%b-%Y")+
  scale_color_manual(name = "Lines", values = cols)+
  theme_tufte()+
  ylab("Price Difference")+
  xlab("Date")+
  ggtitle("Day-to-Day Absolute Price Difference of Exxon Mobil Stock")

cols <- c("Predicted Price Difference" = "red", "Actual Price Difference" = "green")
ggplot(data = Exxon_Regression_PriceDiff) +
  geom_line(mapping = aes(x = Date, y = True_Price_Difference, col = "Actual Price Difference"))+
  geom_line(mapping = aes(x = Date, y = Predicted_Price_Difference, col = "Predicted Price Difference"))+
  scale_x_date(date_breaks = "months", date_labels = "%b-%Y")+
  scale_color_manual(name = "Lines", values = cols)+
  theme_tufte()+
  ylab("Price Difference")+
  xlab("Date")+
  ggtitle("Day-to-Day Price Difference of Exxon Mobil Stock")

```

Regression -- Forecasting -- Predicting Actual Price using Price Difference
```{r}

predict_prices3 <- function(initial_price, sentiment_vector, volume_vector, absolute_difference = FALSE){
  
  Price_Diffs <- c()
  n <- 1 
  while(n <= length(sentiment_vector)){
    price_diff <- round(0.00000004*volume_vector[n] - 1.5*sentiment_vector[n])
    Price_Diffs <- append(Price_Diffs, price_diff, after = length(Price_Diffs))
    n = n + 1
    print("Predicting price differences...")
  }
  
  df <- as.data.frame(Price_Diffs)
  colnames(df) <- c("Predicted_Price_Difference")
  df$Abs_Predicted_Price_Difference <- abs(df$Predicted_Price_Difference)
  
  print("Price difference prediction complete")
  print("Now predicting actual prices")
  
  price_differences <- df$Predicted_Price_Difference
  abs_price_differences <- df$Abs_Predicted_Price_Difference
  
  if (absolute_difference == FALSE){
    
    new_price <- initial_price + price_differences[1]
    predicted_prices <- c(new_price)
    
    for (difference in 2:length(price_differences)){
      new_price <- predicted_prices[difference-1] + price_differences[difference]
      predicted_prices <- append(predicted_prices, new_price, after = length(predicted_prices))
      print("Predicting new prices...")
    }
    
  }
  else{
    new_price <- initial_price + abs_price_differences[1]
    predicted_prices <- c(new_price)
    
    for (difference in 2:length(abs_price_differences)){
      new_price <- predicted_prices[difference-1] + abs_price_differences[difference]
      predicted_prices <- append(predicted_prices, new_price, after = length(predicted_prices))
      print("Predicting new prices...")
    }
  }
  
  print("Done")
  return(predicted_prices)

}

Predicted_Price_3 <- predict_prices3(41.86, Exxon_Regression_PriceDiff$Polarity, Exxon_Regression$Volume)

Exxon_Regression_PriceDiff$Predicted_Price_3 <- Predicted_Price_3

cols <- c("Predicted Price" = "red", "Actual Price" = "green")
ggplot(data = Exxon_Regression_PriceDiff)+
  geom_line(mapping = aes(x = Date, y = Price, col = "Actual Price"))+
  geom_line(mapping = aes(x = Date, y = Predicted_Price_3, col = "Predicted Price"))+
  scale_x_date(date_breaks = "months", date_labels = "%b-%Y")+
  scale_color_manual(name = "Lines", values = cols)+
  theme_tufte()+
  ylab("Stock Price")+
  xlab("Date")+
  ggtitle("Exxon Mobil Stock Price")


```

