---
author: "Trist'n Joseph (trisxcjoseph@gmail.com)"
---

Packages
```{r}
# Package to assist with article scarping 
library(rvest) 

# Packages to assist with text cleaning 
library(stringr)
library(tm)
library(tidytext)

# Packages to assit with polarity calculation
library(rJava)
library(qdap)

# Package to assit with data manipulation & plotting 
library(tidyverse)
library(ggthemes)

# Package to collect financial data 
library(quantmod)

```

getArticles Function
```{r}
# This function takes the company name (or name as will be presented on MarketWatch)
# finds the page with the articles regarding that company, collects the articles
# and returns a dataframe with url, datetime, article title and artile body. 
getArticles <- function(company_name){
  # Link concatenates the company name (or name as presented on MarketWatch) with the remainder of the url
  # to create the url for the page contatining all articles about that input company 
  link <- paste("https://www.marketwatch.com/search?q=", company_name, "&m=Keyword&rpp=15&mp=0&bd=false&rs=false", sep="")   
  
  # Reads the link and finds the page 
  all_articles_page <- read_html(link)
  
  # Collects the text from the page 
  datetime <- all_articles_page%>%
  html_nodes("div.deemphasized span")%>%
  html_text()
  
  # Saves the url of each article 
  urls <- all_articles_page%>%
  html_nodes("div.searchresult a") %>%
  html_attr("href")
    
  articles_webpgs <- data.frame(WebPg = urls, datetime)
  articles_webpgs$WebPg <- as.character(articles_webpgs$WebPg)
  keep <- str_starts(articles_webpgs$WebPg, "http") 
  # For some reason, many urls on this page are inactive or broken & do not contain "http(s)"
  articles_webpgs <- articles_webpgs[keep, ]
  
  titles <- c()
  bodies <- c()
  
  # Loops through each article, collects the title & body 
  for (i in articles_webpgs$WebPg){
    articles_webpgs1 <- read_html(i)
    title <- articles_webpgs1 %>%
      html_node("title") %>%
      html_text()
    titles <- append(titles, title)
    
    articles_webpgs1 <- read_html(i)
    body <- articles_webpgs1 %>%
      html_nodes("p") %>%
      html_text()
    one_body <- paste(body, collapse = " ")
    bodies <- append(bodies, one_body)
  } 
  articles_webpgs$Title <- titles
  articles_webpgs$Body <- bodies 
  return(articles_webpgs)
} 
# Needs updating.

```

Exxon Article Data -- Local File
```{r}
exxon_articles <- read_csv("E:/Projects/Sentiment Analysis/ExxonArticle_2.csv")

# Removes unnecessary columns from the working df
exxon_articles$X1 <- NULL
exxon_articles$Date.X <- NULL
ExxonArticle_2 <- exxon_articles # Renamed to match co-author's file
rm(exxon_articles) # no longer needed

# Renames columns in the working df
names(ExxonArticle_2) <- c("Date", "Article")

ExxonArticle_2 <- ExxonArticle_2 %>%
  arrange(Date)

ExxonArticle_Analysis <- ExxonArticle_2 %>%
  filter(Date >= "2020-03-01" & Date <= "2020-06-20")
rm(ExxonArticle_2) # no longer needed

# Calculates the polarity of each article
Polarity <- polarity(ExxonArticle_Analysis$Article) 

# Adds the polarity to the working df
ExxonArticle_Analysis$Polarity <- Polarity$all$polarity
rm(Polarity) # no longer needed

# Aggregates the polarity by day since each day can have multiple articles & multiple polarities
ExxonArticle_TotalPolarity <- ExxonArticle_Analysis %>%
  group_by(Date) %>%
  summarise(Total_polarity = sum(Polarity, na.rm = T)) %>%
  arrange(Date)

ExxonArticle_TotalPolarity$Date
anyNA(ExxonArticle_TotalPolarity$Total_polarity)

```

Exxon Financial Data
```{r}
startDate = as.Date("2020-03-4")
endDate = as.Date("2020-06-21")

# DF is automatically saved as the ticker symbol
getSymbols("XOM", from =startDate, to = endDate)

# Extracts the price column as a numeric list from an xts object 
XOM_Price <- as.numeric(XOM$XOM.Close)

# Drops dates where no articles where recorded
XOM_Price <- XOM_Price[-c(2,3,18,21,75)] 

# Extracts the volume column as a numeric list from an xts object 
XOM_Volume <- as.numeric(XOM$XOM.Volume)

# Drops dates where no articles where recorded
XOM_Volume <- XOM_Volume[-c(2,3,18,21,75)] 

rm(XOM) # no longer needed

```

Data Preparation for Models
```{r}
# Combine the working df with price and volume lists 
Exxon_Regression <- cbind(ExxonArticle_TotalPolarity, XOM_Price, XOM_Volume)
rm(ExxonArticle_TotalPolarity, XOM_Price, XOM_Volume, ExxonArticle_Analysis) # no longer needed

# Raname columns 
colnames(Exxon_Regression) <- c("Date", "Polarity", "Price", "Volume")

# Create new variable to record observations as 'Positive' if polarity >= 0
Exxon_Regression$Polarity_Sign <- ifelse(Exxon_Regression$Polarity >= 0, "Positive", "Negative")

# Create new variable to find the standard prices of exxon stock
Exxon_Regression <- Exxon_Regression %>%
  mutate(Price_Standard = (Price - mean(Price))/sd(Price))

# Create grouping variable 'Month' which is the recoreded month of each observation
Exxon_Regression$Month <- factor(months(Exxon_Regression$Date), levels = month.name)

cols <- c("Polarity" = "red", "Standardized Price" = "blue")
ggplot(data = Exxon_Regression)+
  geom_line(mapping = aes(x = Date, y = Polarity, col = "Polarity"))+
  geom_line(mapping = aes(x = Date, y = Price_Standard, col = "Standardized Price"))+
  scale_x_date(date_breaks = "months", date_labels = "%b-%Y")+
  scale_color_manual(name = "Lines", values = cols)+
  theme_tufte()+
  ylab("Standardized Values")+
  xlab("Date")+
  ggtitle("Relationship Between Polarity & Standardized Price")

```

Price Predicting Function With Different Weightings
```{r}

predict_prices1 <- function(initial_price, initial_sentiment, sentiment_vector){
  # This function takes an initial price and a list of sentiments (polarities)
  # where the first sentiment is the initial sentiment
  # Positive news as a higher weighting than negative news 
  # The function returns a list of predicted stock prices 
  # Predictions are made based on the recursive relationship defined below 
  
  # Function is optimized using intermediate value theorem and the evaluation metric
  
  stock_price <- c()
  if (initial_price < 0){
    new_stock_price <- initial_price + (0.1*initial_price*initial_sentiment)
  }
  else {
    new_stock_price <- initial_price + (0.105*initial_price*initial_sentiment)
  }
  stock_price <- append(stock_price, new_stock_price, after = length(stock_price))
  for(i in 1:length(sentiment_vector)){
    if(sentiment_vector[i] < 0){
      new_stock_price <- stock_price[i] + (0.1*stock_price[i]*sentiment_vector[i])
      print("Predicting new stock prices...")
    }
    else{
      new_stock_price <- stock_price[i] + (0.105*stock_price[i]*sentiment_vector[i])
      print("Predicting new stock prices...")
    }
    stock_price <- append(stock_price, new_stock_price, after = length(stock_price))
  }
  final_stock_prices <- c(initial_price, stock_price)
  print("Done")
  return(final_stock_prices)
}   
# Constant on sentiment_vector is ideal at 0.1 for negative sentiments & 0.105 for positive sentiments

predicted_prices_1 <- predict_prices1(52.42, -0.07987329, Exxon_Regression$Polarity[2:70])
Exxon_Regression$Predicted_Price_1 <- predicted_prices_1

model1_difference <- Exxon_Regression %>%
  # The average deviation should be minimized for ideal model performance
  select(c(Price, Predicted_Price_1)) %>%
  mutate(Difference = Predicted_Price_1 - Price, Absolute_Difference = abs(Price - Predicted_Price_1)) %>%
  summarize(Avg_Predicted_minus_Actual = mean(Difference), Avg_Absolute_Difference = mean(Absolute_Difference))
model1_difference

model1_difference_by_month <- Exxon_Regression %>%
  # This shows model performance broken down by month
  select(c(Price, Predicted_Price_1, Month)) %>%
  mutate(Difference = Predicted_Price_1 - Price, Absolute_Difference = abs(Price - Predicted_Price_1)) %>%
  group_by(Month) %>%
  summarise(Avg_Predicted_minus_Actual = mean(Difference), Avg_Absolute_Difference = mean(Absolute_Difference))
model1_difference_by_month

cols <- c("Predicted Price" = "red", "Actual Price" = "green")
ggplot(data = Exxon_Regression)+
  geom_line(mapping = aes(x = Date, y = Price, col = "Actual Price"))+
  geom_line(mapping = aes(x = Date, y = Predicted_Price_1, col = "Predicted Price"))+
  scale_x_date(date_breaks = "months", date_labels = "%b-%Y")+
  scale_color_manual(name = "Lines", values = cols) +
  theme_tufte()+
  ylab("Stock Price")+
  xlab("Date")+
  ggtitle("Exxon Mobil Stock Price")

```

Price Predicting Function With Different Weightings & Volume 
```{r}

predict_prices2 <- function(initial_price, initial_sentiment, initial_volume, sentiment_vector, volume_vector){
  # This function takes an initial price, a list of sentiments (polarities), and a list of volumes
  # where the first sentiment is the initial sentiment & first volume is initial volume 
  # Positive news as a higher weighting than negative news 
  # The function returns a list of predicted stock prices 
  # Predictions are made based on the recursive relationship defined below 
  
  # Function is optimized using intermediate value theorem and the evaluation metric 
  
  stock_price <- c()
  if (initial_price < 0){
    new_stock_price <- initial_price + (0.1*initial_price*initial_sentiment) - (0.000000006*initial_volume*initial_sentiment)
  }
  else {
    new_stock_price <- initial_price + (0.105*initial_price*initial_sentiment) - (0.000000006*initial_volume*initial_sentiment)
  }
  stock_price <- append(stock_price, new_stock_price, after = length(stock_price))
  for(i in 1:length(sentiment_vector)){
    if(sentiment_vector[i] < 0){
      new_stock_price <- stock_price[i] + (0.1*stock_price[i]*sentiment_vector[i]) - (0.000000006*volume_vector[i]*sentiment_vector[i])
      print("Predicting new stock prices...")
    }
    else{
      new_stock_price <- stock_price[i] + (0.105*stock_price[i]*sentiment_vector[i]) - (0.000000006*volume_vector[i]*sentiment_vector[i])
      print("Predicting new stock prices...")
    }
    stock_price <- append(stock_price, new_stock_price, after = length(stock_price))
  }
  final_stock_prices <- c(initial_price, stock_price)
  print("Done")
  return(final_stock_prices)
} 
# Constant on volume_vector <= 0.00000001 & >= 0.000000006
# Constant on sentiment_vector is ideal at 0.1 for negative sentiments & 0.105 for positive sentiments
# when constant on volume_vector is within the stated range

predicted_prices_2 <- predict_prices2(52.42, -0.07987329, 24079600, Exxon_Regression$Polarity[2:70], Exxon_Regression$Volume[2:70])
Exxon_Regression$Predicted_Price_2 <- predicted_prices_2

model2_difference <- Exxon_Regression %>%
  # The average deviation should be minimized for ideal model performance
  select(c(Price, Predicted_Price_2)) %>%
  mutate(Difference = Predicted_Price_2 - Price, Absolute_Difference = abs(Price - Predicted_Price_2)) %>%
  summarize(Avg_Predicted_minus_Actual = mean(Difference), Avg_Absolute_Difference = mean(Absolute_Difference))
model2_difference

model2_difference_by_month <- Exxon_Regression %>%
  # This shows model performance broken down by month
  select(c(Price, Predicted_Price_2, Month)) %>%
  mutate(Difference = Predicted_Price_2 - Price, Absolute_Difference = abs(Price - Predicted_Price_2)) %>%
  group_by(Month) %>%
  summarise(Avg_Predicted_minus_Actual = mean(Difference), Avg_Absolute_Difference = mean(Absolute_Difference))
model2_difference_by_month

cols <- c("Predicted Price" = "red", "Actual Price" = "green")
ggplot(data = Exxon_Regression)+
  geom_line(mapping = aes(x = Date, y = Price, col = "Actual Price"))+
  geom_line(mapping = aes(x = Date, y = Predicted_Price_2, col = "Predicted Price"))+
  scale_x_date(date_breaks = "months", date_labels = "%b-%Y")+
  scale_color_manual(name = "Lines", values = cols)+
  theme_tufte()+
  ylab("Stock Price")+
  xlab("Date")+
  ggtitle("Exxon Mobil Stock Price")

```

Price Predicting Function With Different Weightings & Volume (Confidence Interval)
```{r}

create_confidence_intervals <- function(predicted_prices, confidence_level = 0.95){
  # This function creates the confidence intervals for the predicted prices 
  # It accepts a list of prices and an optimal confidence level (default is 95%)
  # Returns a df with confidence interval upper and lower limits 
  
  standard_dev = sd(predicted_prices)
  remainder <- 1 - confidence_level
  confidence_level_probability <- confidence_level + (remainder/2)
  z = qnorm(confidence_level_probability)
  upper_limits <- c()
  for (price in predicted_prices){
    upper_limit <- price + (z*standard_dev/sqrt(length(predicted_prices)))
    upper_limits <- append(upper_limits, upper_limit, after = length(upper_limits))
    print("Calculating upper confidence limits...")
  }
  lower_limits <- c()
  for (price in predicted_prices){
    lower_limit <- price - (z*standard_dev/sqrt(length(predicted_prices)))
    lower_limits <- append(lower_limits, lower_limit, after = length(lower_limits))
    print("Calculating lower confidence limits...")
  }
  df <- cbind(lower_limits, upper_limits)
  df <- as.data.frame(df)
  print("Done")
  return(df)
}

confidence_intervals <- create_confidence_intervals(Exxon_Regression$Predicted_Price_2) # 95% CI
Exxon_Regression_Price_2 <- Exxon_Regression %>%
  select(c("Date", "Price", "Month", "Predicted_Price_2")) %>%
  mutate(ConfInt_lower=confidence_intervals$lower_limits, ConfInt_upper=confidence_intervals$upper_limits)

confidence_interval_80 <- create_confidence_intervals(Exxon_Regression$Predicted_Price_2, 0.80)
Exxon_Regression_Price_2$ConfInt_lower_80 <- confidence_interval_80$lower_limits
Exxon_Regression_Price_2$ConfInt_upper_80 <- confidence_interval_80$upper_limits

confidence_interval_99 <- create_confidence_intervals(Exxon_Regression$Predicted_Price_2, 0.99)
Exxon_Regression_Price_2$ConfInt_lower_99 <- confidence_interval_99$lower_limits
Exxon_Regression_Price_2$ConfInt_upper_99 <- confidence_interval_99$upper_limits

cols <- c(
  "Predicted Price" = "red",
  "Actual Price" = "green", 
  "95% Confidence Interval"="blue",
  "80% Confidence Interval"="orange",
  "99% Confidence Interval"="dark grey"
  )

ggplot(data = Exxon_Regression_Price_2)+
  geom_line(mapping = aes(x = Date, y = Price, col = "Actual Price"))+
  geom_line(mapping = aes(x = Date, y = Predicted_Price_2, col = "Predicted Price"))+
  # 95 % CI
  #geom_line(mapping = aes(x = Date, y = ConfInt_lower, col ="95% Confidence Interval"), linetype = "dashed")+
  #geom_line(mapping = aes(x = Date, y = ConfInt_upper), col="blue", linetype = "dashed")+
  # 80% CI
  #geom_line(mapping = aes(x = Date, y = ConfInt_lower_80, col ="80% Confidence Interval"), linetype = "dashed")+
  #geom_line(mapping = aes(x = Date, y = ConfInt_upper_80), col="orange", linetype = "dashed")+
  # 99% CI
  geom_line(mapping = aes(x = Date, y = ConfInt_lower_99, col ="99% Confidence Interval"), linetype = "dashed")+
  geom_line(mapping = aes(x = Date, y = ConfInt_upper_99), col="dark grey", linetype = "dashed")+
  scale_x_date(date_breaks = "months", date_labels = "%b-%Y")+
  scale_color_manual(name = "Lines", values = cols)+
  theme_tufte()+
  ylab("Stock Price")+
  xlab("Date")+
  ggtitle("Exxon Mobil Stock Price")

```

Predicting Price Difference
```{r}

Exxon_Regression_PriceDiff <- Exxon_Regression[2:nrow(Exxon_Regression), ] %>% 
  select(c(Date, Polarity, Price, Volume, Month))

calculate_darily_price_difference <- function(Prices){
  # This function takes a list of prices
  # Returns a df containing price difference and abs price difference 
  
  row = 1
  Price_Diffs = c()
  while(row < length(Prices)){
    price_diff <- round(Prices[row+1] - Prices[row], 3)
    Price_Diffs <- append(Price_Diffs, price_diff, after = length(Price_Diffs))
    row = row + 1
    print("Calculating price differences...")
  }
  odf <- as.data.frame(Price_Diffs)
  colnames(odf) <- c("True_Price_Difference")
  odf$Abs_Price_Difference <- abs(odf$True_Price_Difference)
  print("Done")
  return(odf)
}  

Actual_PriceDifferences <- calculate_darily_price_difference(Exxon_Regression$Price)
Exxon_Regression_PriceDiff <- cbind(Exxon_Regression_PriceDiff, Actual_PriceDifferences) 

cols <- c("Polarity" = "red", "Absolute Price Difference" = "blue")
ggplot(data = Exxon_Regression_PriceDiff)+
  geom_line(mapping = aes(x = Date, y = Abs_Price_Difference, col = "Absolute Price Difference"))+
  geom_line(mapping = aes(x = Date, y = Polarity, col = "Polarity"))+
  scale_x_date(date_breaks = "months", date_labels = "%b-%Y")+
  scale_color_manual(name = "Lines", values = cols)+
  theme_tufte()+
  ylab("")+
  xlab("Date")+
  ggtitle("Relationship Between Price Difference & Polarity of Exxon Mobil Stock")


predict_price_differences <- function(sentiment_vector, volume_vector){
  # This function takes a list of sentiments (polarities) and volumes
  # Returns a df containing predicted price differences and abs predicted price differences 
  # Predictions are solely based on polarity and traded volume 
  
  Price_Diffs <- c()
  n <- 1 
  while(n <= length(sentiment_vector)){
    price_diff <- round(0.00000004*volume_vector[n] - 1.5*sentiment_vector[n])
    Price_Diffs <- append(Price_Diffs, price_diff, after = length(Price_Diffs))
    n = n + 1
    print("Predicting price differences...")
  }
  odf <- as.data.frame(Price_Diffs)
  colnames(odf) <- c("Predicted_Price_Difference")
  odf$Abs_Predicted_Price_Difference <- abs(odf$Predicted_Price_Difference)
  print("Done")
  return(odf)
}
# Constant on volume_vector <= 0.000000045 & >= 0.000000040
# Constant on sentiment_vector is ideal at 1.5 when constant on volume_vector is within the stated range

predicted_price_difference <- predict_price_differences(Exxon_Regression_PriceDiff$Polarity, Exxon_Regression$Volume)
Exxon_Regression_PriceDiff$Predicted_Price_Difference <- predicted_price_difference$Predicted_Price_Difference
Exxon_Regression_PriceDiff$Abs_Predicted_Price_Difference <- predicted_price_difference$Abs_Predicted_Price_Difference

price_difference_model_performance <- Exxon_Regression_PriceDiff %>% 
  # The average deviation should be minimized for ideal model performance
  select(c(Abs_Price_Difference, Abs_Predicted_Price_Difference)) %>% 
  mutate(Deviation = Abs_Predicted_Price_Difference - Abs_Price_Difference) %>% 
  summarise(Average_Deviation = mean(Deviation))
price_difference_model_performance 
 

price_difference_model_performance_by_month <- Exxon_Regression_PriceDiff %>% 
  # This shows model performance broken down by month
  select(c(Abs_Price_Difference, Abs_Predicted_Price_Difference, Month)) %>% 
  group_by(Month) %>% 
  mutate(Deviation = Abs_Predicted_Price_Difference - Abs_Price_Difference) %>% 
  summarise(Average_Deviation = mean(Deviation))
price_difference_model_performance_by_month

cols <- c("Predicted Absolute Price Difference" = "red", "Actual Absolute Price Difference" = "green")
ggplot(data = Exxon_Regression_PriceDiff) +
  geom_line(mapping = aes(x = Date, y = Abs_Price_Difference, col = "Actual Absolute Price Difference"))+
  geom_line(mapping = aes(x = Date, y = Abs_Predicted_Price_Difference, col = "Predicted Absolute Price Difference"))+
  scale_x_date(date_breaks = "months", date_labels = "%b-%Y")+
  scale_color_manual(name = "Lines", values = cols)+
  theme_tufte()+
  ylab("Price Difference")+
  xlab("Date")+
  ggtitle("Day-to-Day Absolute Price Difference of Exxon Mobil Stock")

cols <- c("Predicted Price Difference" = "red", "Actual Price Difference" = "green")
ggplot(data = Exxon_Regression_PriceDiff) +
  geom_line(mapping = aes(x = Date, y = True_Price_Difference, col = "Actual Price Difference"))+
  geom_line(mapping = aes(x = Date, y = Predicted_Price_Difference, col = "Predicted Price Difference"))+
  scale_x_date(date_breaks = "months", date_labels = "%b-%Y")+
  scale_color_manual(name = "Lines", values = cols)+
  theme_tufte()+
  ylab("Price Difference")+
  xlab("Date")+
  ggtitle("Day-to-Day Price Difference of Exxon Mobil Stock")

```

Predicting Actual Price using Price Difference
```{r}

predict_prices3 <- function(initial_price, sentiment_vector, volume_vector, absolute_difference = FALSE){
  # This function takes an initial price, a list of sentiments (polarities) and volumes, and an optinal absolute difference parameter
  # By default, the function determines the true  
  # Price difference predictions are solely based on polarity and traded volume
  # Predicted prices are determined recursively using the predicted price differences
  # Returns a list of predicted prices
 
  Price_Diffs <- c()
  n <- 1 
  while(n <= length(sentiment_vector)){
    price_diff <- round(0.00000004*volume_vector[n] - 1.5*sentiment_vector[n])
    Price_Diffs <- append(Price_Diffs, price_diff, after = length(Price_Diffs))
    n = n + 1
    print("Predicting price differences...")
  }
  
  df <- as.data.frame(Price_Diffs)
  colnames(df) <- c("Predicted_Price_Difference")
  df$Abs_Predicted_Price_Difference <- abs(df$Predicted_Price_Difference)
  
  print("Price difference prediction complete")
  print("Now predicting actual prices")
  
  price_differences <- df$Predicted_Price_Difference
  abs_price_differences <- df$Abs_Predicted_Price_Difference
  
  if (absolute_difference == FALSE){
    
    new_price <- initial_price + price_differences[1]
    predicted_prices <- c(new_price)
    
    for (difference in 2:length(price_differences)){
      new_price <- predicted_prices[difference-1] + price_differences[difference]
      predicted_prices <- append(predicted_prices, new_price, after = length(predicted_prices))
      print("Predicting new prices...")
    }
    
  }
  else{
    new_price <- initial_price + abs_price_differences[1]
    predicted_prices <- c(new_price)
    
    for (difference in 2:length(abs_price_differences)){
      new_price <- predicted_prices[difference-1] + abs_price_differences[difference]
      predicted_prices <- append(predicted_prices, new_price, after = length(predicted_prices))
      print("Predicting new prices...")
    }
  }
  
  print("Done")
  return(predicted_prices)
}

Predicted_Price_3 <- predict_prices3(41.86, Exxon_Regression_PriceDiff$Polarity, Exxon_Regression$Volume)
Exxon_Regression_PriceDiff$Predicted_Price_3 <- Predicted_Price_3

cols <- c("Predicted Price" = "red", "Actual Price" = "green")
ggplot(data = Exxon_Regression_PriceDiff)+
  geom_line(mapping = aes(x = Date, y = Price, col = "Actual Price"))+
  geom_line(mapping = aes(x = Date, y = Predicted_Price_3, col = "Predicted Price"))+
  scale_x_date(date_breaks = "months", date_labels = "%b-%Y")+
  scale_color_manual(name = "Lines", values = cols)+
  theme_tufte()+
  ylab("Stock Price")+
  xlab("Date")+
  ggtitle("Exxon Mobil Stock Price")

```
